---
title: "Statistical Modelling for Infectious Disease Management - Prediction of future infections in a group"
output:
  rmarkdown::html_vignette:
    fig_width: 7 
    fig_height: 5
fontsize: 12pt
geometry: margin=1in
vignette: >
  %\VignetteIndexEntry{Statistical Modelling for Infectious Disease Management - Prediction of future infections in a group}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE
)

knitr::knit_hooks$set(
  source = function(x, options) {
    hook.r <- function(x, options) {
      fence <- "```"
      language <- tolower(options$engine)
      if (language == 'node') language <- 'javascript'
      if (!options$highlight) language <- 'text'
      if(!is.null(options$foldcode)) {
      paste0('\n\n', "<details><summary>Source</summary>\n", fence, language, '\n', x, fence,  '\n\n', "</details>\n")
      } else {
              paste0('\n\n', fence, language, '\n', x, fence,  '\n\n')
      }
    }
    x <- knitr:::hilight_source(x, 'markdown', options)
    hook.r(
      paste(c(
        x, 
        ''
      ), collapse = '\n'), 
      options
    )
  }
)

Sys.setlocale("LC_TIME", "C")

```

```{r setup, echo = FALSE, message = FALSE}
library(smidm)
library(ggplot2)

```

## Question

A family party with 25 participants took place on the 15.03.2022 in a closed room and no masks were worn. In the following days some participants start to show symptoms of a COVID-19 infection: one on 18.03.2022 and three on 20.03.2022. How many further symptomatic infections can be expected in the following days?

## Calculating a prediction of the total number of infections with `get_expected_total_infections`

The function `get_expected_total_infections()` can be used to give a first answer to the question. It returns a prediction how many people in the group are expected to show symptoms in total.

### Inputs

The following input values are necessary for the function `get_expected_total_infections()` to work:

The `group_size` is the number of people participating in the event, including all observed infections.

The `last_day_reported_infection` is the number of days after the event when the last symptom begin was observed and the `total_reported_infections` is the total number of observed infections so far. 

Finally, `meanlog` and `sdlog` are the mean and standard deviation parameters of the log-normal distribution for the incubation time derived from the paper Xin et al. [1].

### Methodology

Based on the incubation time distribution one can calculate the percentage of all symptomatic infections that will have their symptom onset up to the `last_day_reported_infection`. Then, said percentage is combined with the `total_reported_infections` to calculate the total symptomatic infections. The minimum between the result and the `group_size` is returned, because the `group_size`is obviously an upper bound for the total infections.

### Output

```{r get_expected_total_infections}
group_size <- 25
last_day_reported_infections <- 5 # day 0 = event day
total_reported_infections <- 4
meanlog <- 1.69
sdlog <- 0.55

predicted_total_infections <- get_expected_total_infections(group_size,
                                                            last_day_reported_infections,
                                                            total_reported_infections,
                                                            meanlog,
                                                            sdlog)
print(predicted_total_infections)

```

The output represents how many people are expected to get a symptomatic infection.
In the example 10 infections are predicted in total, which implies one can expect 6 further people starting to show symptoms in the next days because 4 infections were already observed.

## Generating a vector with number of people starting to show symptoms on each day using \newline `predict_future_infections`

The function `predict_future_infections()` can be used to give a more detailed answer.
It creates a vector containing the predicted number of further people starting to show symptoms on each of the days after the event.

### Inputs

Multiple arguments are necessary for the function `predict_future_infections()` to work:

The `last_day_reported_infection` is the number of days after the event when the last symptom begin was observed and the `total_reported_infections` is the total number of observed infections so far. 

Then, the `total_expected_infections` is needed, which defines the total number of expected infections, including the ones already observed. One can use the output of `get_expected_total_infections()` or an own estimation based on e.g. reported symptomatic infection rates in a population of interest. If the output of `get_expected_total_infections()` is used, then it should be based on the same `meanlog` and `sdlog` as in the call to `predict_future_infections()`.

Finally, `meanlog` and `sdlog` are the mean and standard deviation parameters of the log-normal distribution for the incubation time.

### Methodology

The function `predict_future_infections()` uses the function `get_incubation_day_distribution()` to get a vector of day-specific probabilities of symptom onset, given that a person will develop symptoms. Default values of the log-normal distribution for the incubation time used in that function are taken from the paper Xin et al. [1].

Starting on the first day after the `last_day_reported_infection`, the probability of symptom onset on a particular day, given that no symptoms occurred so far, is multiplied by the number of further expected infections and rounded upwards to receive the expected number of people starting to show symptoms on that day. 
The probability that no symptoms occurred so far can be calculated by 1 minus the sum of symptom onset probabilities for all previous days. 
The number of further expected infections is simply `total_expected_infections` minus `total_reported_infections`. 
The latter is afterwards raised by the predicted number of people with symptom onset on the day currently looked at, so that afterwards the next day can be treated.

When at some point the updated `total_reported_infections` is not smaller than `total_expected_infections` anymore, a 0 is inserted to signal that all further expected symptomatic infections are allocated and the loop is stopped.
 
### Output

```{r predict_future_infections}
last_day_reported_infections <- 5 # day 0 = event day
total_reported_infections <- 4
total_expected_infections <- get_expected_total_infections(25, 5, 4)
meanlog <- 1.69
sdlog <- 0.55

predicted_daily_infections <- predict_future_infections(last_day_reported_infections,
                                                        total_reported_infections,
                                                        total_expected_infections,
                                                        meanlog,
                                                        sdlog)
print(predicted_daily_infections)

```

The function `predict_future_infections()` creates a vector with values representing the expected distribution of new symptomatic infections over the days after the event. Up to the `last_day_reported_infections` the entries are 0 because only infections in the future are predicted. 

## An example for visualizing the output of \newline `predict_future_infections` 

```{r libraries, foldcode = TRUE, message = FALSE}
data <- data.frame("Erkrankungsdatum" = as.Date("2022-03-15") + 0:5,
                   "Neue_Faelle" = c(0, 0, 0, 1, 0, 3))
expected <- data.frame("Erkrankungsdatum" = as.Date("2022-03-15") + 1:(length(predicted_daily_infections)),
                       "ErwarteteWeitereFaelle" = predicted_daily_infections)
g <- ggplot(expected) +
    geom_bar(
      data = data,
      aes(Erkrankungsdatum,
          Neue_Faelle,
          fill = "observation"),
      stat = 'identity'
    ) +
    geom_bar(
      data = expected,
      aes(Erkrankungsdatum,
          ErwarteteWeitereFaelle,
          fill = "prediction"
      ),
      stat = 'identity'
    ) +
    geom_vline(xintercept = expected$Erkrankungsdatum[1]) +
    geom_label(aes(x = expected$Erkrankungsdatum[1], y = data$Neue_Faelle[1] + 4, label = "event"), 
               colour = "black", fill = "white", vjust = 1, size = 7) +
    scale_y_continuous(breaks = function(x) unique(
      floor(pretty(seq(0, (max(x) + 1) * 1.1))))) +
    scale_x_date(date_breaks = "1 day", date_labels = "%d %b", minor_breaks = NULL) +
    ylab("infected") +
    xlab("timeline") +
    labs(fill = 'type of cases') +
    theme(legend.position = c(0.75,0.85), text = element_text(size = 16),
          axis.text.x = element_text(face = "bold", angle = 30, hjust = 1))
  return(g)


```

## Literature

[1] Xin H, Wong JY, Murphy C, Yeung A, Ali ST, Wu P, Cowling BJ. The Incubation Period Distribution of Coronavirus Disease 2019: A Systematic Review and Meta-Analysis. Clinical Infectious Diseases. 2021; 73(12): 2344-2352.
