---
title: "Statistical Modelling for Infectious Disease Management - Infection period"
output:
  rmarkdown::html_vignette:
    fig_width: 7 
    fig_height: 5
fontsize: 12pt
geometry: margin=1in
vignette: >
  %\VignetteIndexEntry{Statistical Modelling for Infectious Disease Management - Infection period}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE
)

knitr::knit_hooks$set(
  source = function(x, options) {
    hook.r <- function(x, options) {
      fence <- "```"
      language <- tolower(options$engine)
      if (language == 'node') language <- 'javascript'
      if (!options$highlight) language <- 'text'
      if(!is.null(options$foldcode)) {
      paste0('\n\n', "<details><summary>Source</summary>\n", fence, language, '\n', x, fence,  '\n\n', "</details>\n")
      } else {
              paste0('\n\n', fence, language, '\n', x, fence,  '\n\n')
      }
    }
    x <- knitr:::hilight_source(x, 'markdown', options)
    hook.r(
      paste(c(
        x, 
        ''
      ), collapse = '\n'), 
      options
    )
  }
)

Sys.setlocale("LC_TIME", "C")

```

```{r setup, echo = FALSE, message = FALSE}
library(smidm)
library(ggplot2)
library(dplyr)
library(hdrcde)

```

## Question

One or several persons start to show symptoms of COVID-19. When did the person become infected?

## Generating a data frame with dates and infection probability using \newline `get_infection_density` for one person

The function `get_infection_density()` can be used to calculate a data frame containing the infection probability when the person shows symptoms.

### Inputs

The function `get_infection_density()` expects the following input arguments:

First, the `symptom_begin_date` is needed, which defines when the person started to have symptoms. 

Then, the `max_incubation_days` has to be set, which defines the interval length of the distribution output. 

The remaining inputs `meanlog` and `sdlog` are the parameters of the log-normal distribution for the infection probability.

```{r get_infection_date_density_statement}
symptom_begin_date <- as.Date("2021-12-28")
max_incubation_days <- 18
meanlog <- 1.69
sdlog <- 0.55

infec_date_df <- get_infection_density(symptom_begin_date,
                                       max_incubation_days,
                                       meanlog,
                                       sdlog)

```

### Methodology

The default values of log-normal distribution are taken from the paper Xin et al [1]. In this paper the authors made a systematic review of the current literature and estimated those parameters based on their meta-analysis.  

### Output

The data frame shows for each hour from the earliest potential start of infection up to the symptom begin date the resulting density of the log-normal distribution. This density can be used for calculating the most probable period of the infection.

```{r get_infection_density_result, echo = FALSE, message = FALSE}
knitr::kable(infec_date_df[100:109, ],
             caption = "values 100 to 109 of resulting data frame")

```

## Generating a data frame with dates and probability of infection using \newline `get_misc_infection_density` for several persons

The function `get_misc_infection_density()` creates a data frame containing the mixture probability of all considered persons. It can be used to give an overview of the infection probability of several persons with symptom onset dates, e.g., one person with symptom onset on 24.12.2021 and two persons with symptom onset on 28.12.2021.

### Inputs

The following arguments are needed for using the function `get_misc_infection_density()`:

The first parameter `symptom_begin_dates` contains the dates when the persons got symptoms. 

The second parameter `persons` contains the number of persons having symptoms on each date. 

The remaining inputs are the same as in `get_infection_density`.

```{r get_misc_infection_date_density}
symptom_begin_dates <- c(as.Date("2021-12-24"), as.Date("2021-12-28"))
persons <- c(1, 2)
max_incubation_days <- 18

misc_infec_date_df <- get_misc_infection_density(symptom_begin_dates,
                                                 persons,
                                                 max_incubation_days)

```

### Methodology

This function uses the `get_infection_density` function and generates a mixture distribution [2]. This probability distribution is obtained by a sum of the infection probability distribution for each symptom onset day multiplied by the percentage of persons, which have started to show symptoms on this day.

### Output

The data shows the mixture log-normal distribution and thus gives an overview of the potential infection time points for all considered persons. However, it does not necessarily have to imply that they had their infection on the same time point. In fact, there did not have to be an event, where the persons met. It shows when the persons got infected and it is possible that there is more than one infection date, which can be seen based on several maxima.  

```{r get_mic_infection_density_result, echo = FALSE, message = FALSE}
knitr::kable(misc_infec_date_df[100:109, ],
             caption = "values 100 to 109 of resulting data frame")

```

## Visualization example of the data frame of \newline `get_infection_density`

```{r .calculate_qstart_qend, foldcode = TRUE}
.calculate_qstart_qend <- function(probability, df) {
    hdr_df <- hdr(den = data.frame(x = 1:length(df$distribution), y = df$distribution),
                  p = probability * 100)$hdr
    qstart <- (hdr_df[1:(length(hdr_df) / 2) * 2] - 1) / 24
    qend  <- (hdr_df[1:(length(hdr_df) / 2) * 2 - 1] - 1) / 24
  return(list("qstart" = qstart, "qend" = qend))
}

```

```{r .shade_curve, foldcode = TRUE}
.shade_curve <- function(df, qstart, qend, fill = "red", alpha = 0.4) {
  subset_df <- df[floor(qstart * 24):ceiling(qend * 24), ]
  geom_area(data = subset_df,
            aes(x = x, y = y), 
            fill = fill, 
            color = NA, 
            alpha = alpha)
}

```

```{r parameters for visualization of get_infection_density, foldcode = TRUE}
  symptom_begin_date <- as.Date("2021-12-28")

  df <- infec_date_df
  period_80 <- .calculate_qstart_qend(0.8, df) 
  period_95 <- .calculate_qstart_qend(0.95, df)
  
  symp_date_posixct_start <- as.POSIXct(format(as.POSIXct(symptom_begin_date, tz = "CET"), "%Y-%m-%d"))
  symp_date_posixct_end <- as.POSIXct(format(as.POSIXct(symptom_begin_date + 1, tz = "CET"), "%Y-%m-%d"))
  symp_date_posixct_mid <- symp_date_posixct_start - as.numeric(difftime(symp_date_posixct_start,
                           symp_date_posixct_end, units = "hours")) / 2 * 3600 
```
  
```{r visualization of get_infection_density, foldcode = TRUE}  
  g <- ggplot() +

    scale_x_datetime(breaks = scales::date_breaks("1 days"), labels = scales::date_format("%d %b")) +
    theme(axis.text.x = element_text(angle = 90)) +
    # scale_x_continuous(breaks = x_tick,
    #                    labels = x_label) +
    # theme(axis.ticks.x = element_line(color = c(rbind(rep("black", length(x_label) / 2), rep(NA, length(x_label) / 2))),        linetype = 2, size = 1))+
    geom_path(aes(x = df$dates, y = df$distribution, color = "red")) +
    .shade_curve(df = data.frame(x = df$dates, y = df$distribution),
                 period_80$qstart,
                 period_80$qend) +
    .shade_curve(df = data.frame(x = df$dates, y = df$distribution),
                 period_95$qstart,
                 period_95$qend,
                 alpha = 0.2) +
    geom_rect(data = data.frame(xmin = symp_date_posixct_start,
                            xmax = symp_date_posixct_end,
                            ymin = -Inf,
                            ymax = Inf),
          aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
          fill = "brown", alpha = 0.3) +
    geom_label(aes(x = symp_date_posixct_mid, y = 0.9*max(df$distribution), label = "symptom\nonset"),
              colour = "brown", fill = "white", size = 5, label.size = NA) +
    ylab("probability") +
    xlab("timeline") +
    labs(color = 'Verteilung') +
    # ggtitle("Visualization of get_infection_density ") +
    theme(legend.position = "none", text = element_text(size = 16*5/5)) +
    theme(axis.text.x = element_text(colour = "black", face = "bold", angle = 30, hjust = 1)) +
    theme(axis.title.x = element_text(colour = "black", face = "bold")) +
    theme(axis.text.y = element_text(colour = "gray50")) +
    theme(axis.title.y = element_text(colour = "gray50"))


  g
  
```

## Visualization example of the data frame of \newline `get_misc_infection_density`

```{r parameters for visualization of get_misc_infection_density, foldcode = TRUE}
  df <- misc_infec_date_df
  
  period_80 <- .calculate_qstart_qend(0.8, df) 
  period_95 <- .calculate_qstart_qend(0.95, df)
  
  symp_date_posixct_start <- as.POSIXct(format(as.POSIXct(symptom_begin_date, tz = "CET"), "%Y-%m-%d"))
  symp_date_posixct_end <- as.POSIXct(format(as.POSIXct(symptom_begin_date + 1, tz = "CET"), "%Y-%m-%d"))
  symp_date_posixct_mid <- symp_date_posixct_start - as.numeric(difftime(symp_date_posixct_start,
                           symp_date_posixct_end, units = "hours")) / 2 * 3600 
  
```
  
```{r visualization of get_misc_infection_density, foldcode = TRUE}  
  g <- ggplot() +

    scale_x_datetime(breaks = scales::date_breaks("1 days"), labels = scales::date_format("%d %b")) +
    theme(axis.text.x = element_text(angle = 90)) +
    # scale_x_continuous(breaks = x_tick,
    #                    labels = x_label) +
    # theme(axis.ticks.x = element_line(color = c(rbind(rep("black", length(x_label) / 2), rep(NA, length(x_label) / 2))),        linetype = 2, size = 1))+
    geom_path(aes(x = df$dates, y = df$distribution, color = "red")) +
    .shade_curve(df = data.frame(x = df$dates, y = df$distribution),
                 period_80$qstart,
                 period_80$qend) +
    .shade_curve(df = data.frame(x = df$dates, y = df$distribution),
                 period_95$qstart,
                 period_95$qend,
                 alpha = 0.2) +
    ylab("probability") +
    xlab("timeline") +
    labs(color = 'Verteilung') +
    # ggtitle("Visualization of get_infection_density") +
    theme(legend.position = "none", text = element_text(size = 16 * 5 / 5)) +
    theme(axis.text.x = element_text(colour = "black", face = "bold", angle = 30, hjust = 1)) +
    theme(axis.title.x = element_text(colour = "black", face = "bold")) +
    theme(axis.text.y = element_text(colour = "gray50")) +
    theme(axis.title.y = element_text(colour = "gray50"))

  g
  
```

## Literature

[1] Xin H, Wong JY, Murphy C, Yeung A, Taslim Ali S, Wu P, Cowling BJ. The Incubation Period Distribution of Coronavirus Disease 2019: A Systematic Review and Meta-Analysis. Clinical Infectious Diseases, 2021; 73(12): 2344-2352.

[2] https://en.wikipedia.org/wiki/Mixture_distribution
